Solution of Differential Algebraic Equations with Local Constraints on the Latent Space of an Autoencoder

### 2019-08-23 paper/paper.tex
Updated the paper; added the abstract I submitted to agu

\title{Solution of Differential Algebraic Equations with Local Constraints
on the Latent Space of an Autoencoder}
\author[1]{Alejandro Francisco Queiruga*}
\address[1]{
  \orgdiv{Energy Geosciences Division}, \orgname{Lawrence Berkeley
    National Laboratory}, \orgaddress{\state{CA}}, \country{USA}}
\corres{*\email{afqueiruga@lbl.gov}}

@@

@@

\abstract[Summary]{
  Multiphase simulation is hard.
Our proposed method is motivated by the problems of multiphase reaction and
transport, where the underlying constraints are the ill-behaved material
properties which are difficult to describe.
Our proposed approach is to build upon modern machine learning
techniques to aid in equation discovery. An autoencoder is trained on
the original material dataset to learn a lower-dimensional
representation of the underlying manifold. The model is designed to be
directly incorporated into a macroscale simulation of physical balance
laws, where now the traditional partial differential equations are being
solved on the learned latent space.
We train an autoencoder to determine the lower-dimensional
representation of the EOS surface using modern deep learning techniques
without using any phase labels. The decoder end of the model is then
plugged back into the balance laws, eliminating the constraint equation.
This provides a more robust
numerical algorithm, while putting the onus of programming part of the
simulation itself onto the computer: tens of thousands of lines of human
written code consisting of theoretical relations, empirical fits, and branching logic are
replaced by code generated by the learned model. The methodology will
first be demonstrated on datasets based on thermodynamic intensive
variables, which are already challenging to analyze, before being
extended to richer datasets such as experimental images or molecular
dynamics. We stress that the proposed approach is only utilizing machine
learning on material data: the simulation is indeed solving physical
balance laws. Through deep learning approaches combined with
domain-specific goals, we can learn representations that are both
interpretable and directly apply to analyzing larger systems.
We illustrate the methodology on two traditional problems: 1) the
pendulum, and 2) the continuous stirred tank reactor.}

@@

\begin{algorithm}
  \caption{\label{alg:statemachine}Statemachine transition logic}
\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{switch}\NormalTok{ phase_old:}
  \ControlFlowTok{case}\NormalTok{ gas:}
    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
\NormalTok{      phase_new = liquid_gas}
  \ControlFlowTok{case}\NormalTok{ liquid:}
    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
\NormalTok{      phase_new = liquid_gas}
  \ControlFlowTok{case}\NormalTok{ liquid_gas:}
    \ControlFlowTok{if}\NormalTok{ S_gas >= }\DecValTok{1}\NormalTok{:}
\NormalTok{      phase_new = gas}
    \ControlFlowTok{if}\NormalTok{ S_liquid >= }\DecValTok{1}\NormalTok{:}
\NormalTok{      phase_new = gas}
\end{Highlighting}
\end{Shaded}
\end{algorithm}

@@

\section*{Acknowlegements}

>>>>

\title{Solution of Differential Algebraic Equations with Local Constraints
on the Latent Space of an Autoencoder}
\author[1]{Alejandro Francisco Queiruga*}
\address[1]{\orgdiv{Energy Geosciences Division}, \orgname{Lawrence Berkeley
    National Laboratory}, \orgaddress{\state{CA}}, \country{USA}}
\corres{*\email{afqueiruga@lbl.gov}}

@@

\abstract[Abstract]{
  Multiphase simulation is hard.
Our proposed method is motivated by the problems of multiphase reaction and
transport, where the underlying constraints are the ill-behaved material
properties which are difficult to describe.
Our proposed approach is to build upon modern machine learning
techniques to aid in equation discovery. An autoencoder is trained on
the original material dataset to learn a lower-dimensional
representation of the underlying manifold. The model is designed to be
directly incorporated into a macroscale simulation of physical balance
laws, where now the traditional partial differential equations are being
solved on the learned latent space.
We train an autoencoder to determine the lower-dimensional
representation of the EOS surface using modern deep learning techniques
without using any phase labels. The decoder end of the model is then
plugged back into the balance laws, eliminating the constraint equation.
This provides a more robust
numerical algorithm, while putting the onus of programming part of the
simulation itself onto the computer: tens of thousands of lines of human
written code consisting of theoretical relations, empirical fits, and branching logic are
replaced by code generated by the learned model. The methodology will
first be demonstrated on datasets based on thermodynamic intensive
variables, which are already challenging to analyze, before being
extended to richer datasets such as experimental images or molecular
dynamics. We stress that the proposed approach is only utilizing machine
learning on material data: the simulation is indeed solving physical
balance laws. Through deep learning approaches combined with
domain-specific goals, we can learn representations that are both
interpretable and directly apply to analyzing larger systems.
We illustrate the methodology on two traditional problems: 1) the
pendulum, and 2) the continuous stirred tank reactor.}

@@

\keywords{Class file; \LaTeXe; \emph{Wiley NJD}}

@@

\begin{algorithm}
\begin{lstlisting}[caption={Finite State Machine transition logic}, label={alg:statemachine}]
switch phase_old:
  case gas:
    if p,T crossing boundary:
      phase_new = liquid_gas
  case liquid:
    if p,T crossing boundary:
      phase_new = liquid_gas
  case liquid_gas:
    if S_gas >= 1:
      phase_new = gas
    if S_liquid >= 1:
      phase_new = gas
\end{lstlisting}
\end{algorithm}
%  \begin{verbatim}

%\end{verbatim}

@@

\section*{Acknowlegements}
%\acks
